\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2024-11-05 Manuscript/reply.tex
updated responses to comments. Will go back through after replies to email

\newcommand{\jy}[1]{\textcolor{red}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
\newcommand{\mc}[1]{\textcolor{green}{(MC: #1)}}

@@

The manuscript has been revised accordingly with the most notable change being 
the addition of Section 5 (Classroom Implementation). We believe this revision 
has enhanced the both the quality and utility of this paper.

@@

\jy{What is ``the first point?''  Please quote everything from the
  reviewer, including the positive comments.}

\begin{comment}
To address the first point, the paper relies on descriptive statistics, ranked
tests, and the results of a set of GLMMs. The descriptive data analysis lends
credence to the authors' argument that the reaction times at the 2022 World
Championships was faster than at any of the relevant comparisons. The rank test
appears to be set up correctly and I believe the authors are interpreting the
results correctly. Similarly, I believe the GLMM results are arrived at
correctly given the provided code.

@@

Thank you for these suggestions, we were able to expand the scope of our data
by including results from the men's and women's 100 meter dash.  We considered
adding additional sprinting events such as the 200 meter dash but
when we included a term for type of event in our model the coefficient on this
event term was significant for the 200 meter dash for both men's and women's
data. The coefficient was not signficant when we included the 100 meter dash
results. This is not particularly suprising as athletes who compete in the longer
events may not have as good of a start as those in the 100
meter events because reaction time plays a much smaller role in the outcome of
the race. While it is common for amateur and collegiate athletes to compete in
multiple sprinting events, it seems reasonable that a world championship 100
meter runner may be only a very good 200 meter runner.  We are trying to
estimate the probabilty of observing a extreme reaction time to determine
whether 0.1 seconds is fair but if we include data from competitions with slower
reaction times, then 0.1 seconds is going to be more extreme.  To answer later
thoughts about looking at all sprinting events, we observed boxplots and what we
found were that reaction times in longer sprinting events were much more stable
than in the 100 meter dash and hurdle and thus any insights would likely not be
informative to questioning whether 0.1 seconds is a fair barrier or if athletes
in those events had faster reaction times in 2022.  This follows the conventional
wisdom that athletes are more focuesed on not false starting at the start of the
longer races than they are on getting a perfect start.

@@

The second reviewer shared a similar sentiment about not effectively
presenting ideas so we we made improvements to the introduction section to try
and more clearly present our ideas.  We shortened the literature review to only
cite papers that were actually of interest to the paper and tried to clearly
define the two purposes of the paper: to decide of 0.1 seconds is fair, and to
\jy{The order is reversed in the paper.}
identify inconsistensies with the 2022 reaction times. We try to be precise in
our language as we do not have enough evidence to conclude that the sensors were
calibrated differently, or that Seiko (who make the timing devices) is at fault
for the 2022 reaction times being inexplicably faster. Thus we make general
recommendations for World Athletics and note that the reaction times in 2022
were unusually fast with no clear reason.

@@

Thank you for these encouraging words. \jy{State what expansions we
  have done.}

@@

\begin{comment}
I am also not convinced that the GLMM methods used for evaluating the
probability of disqualification are valid. The tests use data of reaction times
from performances that were NOT disqualified.  The data is therefore truncated
with no samples from the “left” or “disqualified” tail. This means that any
conclusions about the probability of a reaction time falling below this
threshold is purely extrapolation from the distributional assumptions. The
authors use a gamma distribution based on its fit to the data they have. In my
opinion, the choice of a gamma distribution is not sufficiently justified to
make extrapolating claims about the probability mass in the upsampled tail. I am
not convinced there is enough data to make a strong argument here, especially
when the previous literature on reaction time uses different distributions.
\end{comment}
Athletes may be disqualified based on PED use, and athletes may not finish a race
due to injury, thus we did not consider these times.  However, it does seem
reasonable to include times from people who were disqualified based on their
reaction time.

@@

\begin{comment}
We have taken additional steps to streamline the manuscript and enhance the
presentation of our work. Thank you for your time and we appreciate re-considering
our submission.
\end{comment}

>>>>

\newcommand{\jy}[1]{\textcolor{red}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
\newcommand{\of}[1]{\textcolor{green}{(MC: #1)}}

@@

The manuscript has been revised accordingly with the most notable change being 
the addition of new data in both the GLMM and rank comparison sections in an
attempt to expand the scope of the analysis and derive beter insights.

@@

\jy{What is ``the first point?''  Please quote everything from the
  reviewer, including the positive comments.}
\of{resolved}

\begin{comment}
Using Devon Allen's disqualification at the 2022 World Track and Field
championships as a hook, this paper poses two questions. First, were reaction
times at the 2022 World Track and Field Championships faster than they had been
earlier in 2022 national-level competitions, the 2019 World Track and Field
Championships, and the 2023 World Track and Field Championships? Second, based
on modeled reaction times, is the current standard of a 0.1 second threshold
for reaction times before disqualification too conservative? To
address the first point, the paper relies on descriptive statistics, ranked
tests, and the results of a set of GLMMs. The descriptive data analysis lends
credence to the authors' argument that the reaction times at the 2022 World
Championships was faster than at any of the relevant comparisons. The rank test
appears to be set up correctly and I believe the authors are interpreting the
results correctly. Similarly, I believe the GLMM results are arrived at
correctly given the provided code.

@@

Thank you for these suggestions, we were able to expand the scope of our data
by including results from the men's and women's 100 meter dash.  We considered
adding additional sprinting events such as the 200 meter dash but
when we included a term for type of event in our model the coefficient on this
event term was not significant for when we included the 100 meter dash
but became significant after we added the 200 meter dash.
This is not particularly suprising as athletes who compete in the longer
events may not have as good of a start as those in the 100
meter events because reaction time plays a much smaller role in the outcome of
the race. While it is common for amateur and collegiate athletes to compete in
multiple sprinting events, it seems reasonable that a world championship 100
meter runner may be only a very good 200 meter runner.  We are trying to
estimate the probabilty of observing a extreme reaction time to determine
whether 0.1 seconds is fair but if we include data from competitions with slower
reaction times, then 0.1 seconds is going to be more extreme.  To answer later
thoughts about looking at all sprinting events, we observed boxplots and what we
found were that reaction times in longer sprinting events were much more stable
than in the 100 meter dash and hurdle and thus any insights would likely not be
informative to questioning whether 0.1 seconds is a fair barrier or if athletes
in those events had faster reaction times in 2022.  This follows the conventional
wisdom that athletes are more focuesed on not false starting at the start of the
longer races than they are on getting a perfect start.

@@

The second reviewer shared a similar sentiment about not effectively
presenting ideas so we we made improvements to the introduction section to try
and more clearly present our ideas.  We shortened the literature review to only
cite papers that were actually of interest to the paper and tried to clearly
define the two purposes of the paper: to identify inconsistensies with the 2022
reaction times and to decide of 0.1 seconds is a fair barrier. We try to be precise in
our language as we do not have enough evidence to conclude that the sensors were
calibrated differently, or that Seiko (who make the timing devices) is at fault
for the 2022 reaction times being inexplicably faster. Thus we make general
recommendations for World Athletics and note that the reaction times in 2022
were unusually fast with no clear reason.

@@

Thank you for these encouraging words, we have added additional data to both
the GLMM and clusrank section in order to strengthen our arguments: there was
an anomaly in 2022 and that the reaction time barrier of 0.1 seconds is not
grounded in formal statistical analysis.

\jy{State what expansions we
  have done.}
\of{After we are wrapped up with the manuscript it will be easier to identify
all the improvements we made. Let's just leave this here for now.}

@@

\begin{comment}
I am also not convinced that the GLMM methods used for evaluating the
probability of disqualification are valid. The tests use data of reaction times
from performances that were NOT disqualified.  The data is therefore truncated
with no samples from the “left” or “disqualified” tail. This means that any
conclusions about the probability of a reaction time falling below this
threshold is purely extrapolation from the distributional assumptions. The
authors use a gamma distribution based on its fit to the data they have. In my
opinion, the choice of a gamma distribution is not sufficiently justified to
make extrapolating claims about the probability mass in the upsampled tail. I am
not convinced there is enough data to make a strong argument here, especially
when the previous literature on reaction time uses different distributions.
\end{comment}
We will start by addressing the comments related to runners who have been
disqualified.  Some of the runners who were disqualified had a negative reaction
time, indicating they were trying to predict the starting gun and actually left
before the gun even fired.  These runners cannot be included in the analysis
because the point of the GLMM section is to question the current reaction barrier
methods by determining how fast athletes can react in; but athletes with a
negative time left before the gun even went off (Also practically speaking
a negative reaction time makes little sense).  As for the roughly 20 athletes
across men's and women's data without negative times but were disqualified, we
analyzed how the inclusion of these athletes effected our results.  What we found
was that the probability of a reaction time below 0.1 seconds increased from
0.0029 to 0.0039 and we have attached the R code so that you can see our work.
\of{Not sure the technical details of how this is going to work but we will come
back to this before we submit.} Additionally there were similar parameter values
for the standard deviations of the random effects and we decided that we were
going to keep these observations in our data.

As for why the particular model was
chosen, we looked at several types of models and chose the one that we thought
fit the residual qq plot the best and had the largest log likelihood. After
selecting the gamma generalized linear model we again looked at several link
functions such as "identity", "log",

\of{Need to justify why we did what we did and disadvantages of including the
disqualified
Do this in the paper and reply. Say section where we made any changes. Give
justifications for any reasons for disagreeing}

\of{Will clean this up based on Prof Yan and Schifano's comments to 11/5 email}

@@

We have taken additional steps to streamline the manuscript and enhance the
presentation of our work. Thank you for your time and we appreciate re-considering
our submission.
